Measuring causal influence with\\ back-to-back regression: the linear case - supplementary material

### 2019-05-23 paper/supplementary.tex
new theorem

>>>>

\clearpage
\newpage
A stronger theorem can be proven, that includes provision for noise

THEOREM

    Consider the B2B model from Equation~\ref{eq:model}.
    %
    Assume that $F$ and $X$ are full-rank on the subspace spanned by active entries in $E$.
    %
    Then, the solution of B2B, $\hat H$, will minimise 
    %
    $\min_H  \left \| X - XEH\right\| ^2  + \left \| NEH\right \| ^2$

PROOF

Let $\hat G$ and $\hat H$ be the solutions of the first and second step of B2B, $\hat H=EF\hat G$ and $\hat G$ minimises 
\begin{equation}
	\left \| X - XEFG\right\| ^2  + \left \| NFG\right \| ^2
    \label{eq:doublenorm}
\end{equation}

Let us prove that $EF\hat G = F\hat G$, that is, that the $d_x$ bottom rows of $F\hat G$ are zero.

Let $Z=F^\dagger EF\hat G$, we have $FZ = FF^\dagger EF  \hat G= EF\hat G$ since F is full rank over E. and since E is a contraction, we have 
$$\left \| X - XEFZ\right \| ^2  + \left \| NFZ\right \| ^2 = \left \| X - XEF\hat G\right \| ^2  + \left \| NEF\hat G\right \| ^2 \leq \left \| X - XEF\hat G\right \| ^2  + \left \| NF\hat G\right \| ^2$$

Since $\hat G$ attains the minimum value of \ref{eq:doublenorm}, we must have $EF\hat G = F\Hat G$.

Therefore, $\left \| X - XEFG\right \| ^2  + \left \| NEFG\right \| ^2 = \left \| X - XH\right \| ^2  + \left \| NH\right \| ^2$ are minimised by $\hat G$ and $\hat H$. Noting that $\hat H = E\hat H$ completes the proof.

The solution of $\min_H  \left \| X - XEH\right\| ^2  + \left \| NEH\right \| ^2$ is $\hat H= (E X^\top XE +EN^\top NE) ^\dagger EXX^\top$.

If the covariance matrices of $NE$ and $XE$ (the restriction of N and X to the active causal features) can be co-diagonalised, with orthogonal transformation $U$, then the diagonal elements of $\hat H$ will be those of $U^\top  \frac{\sigma_X}{\sigma_X + \sigma_N} U $.

In any case, the $d_x-k$ last diagonal elements will be zero. All others will be bounded by the maximum and minimum value of $\frac{\sigma_X}{\sigma_X + \sigma_N}$ over all eigenvalues of the covariance matrices of $XE$ and $NE$. If the eigenvalues are sorted from largest to smallest, diagonal elements of $\hat H$ will be between $\frac{\sigma_{X_1}}{\sigma_{X_1} +\sigma_{N_k}}$ and $\frac{\sigma_{X_k}}{\sigma_{X_k} +\sigma_{N_1}}$.
